---
title: "Homework 3"
author: Qianyi Sun
date: March 19, 2020
output: github_document
---



```{r}
library('MASS') ## for 'mcycle'
library('manipulate') ## for 'manipulate'
```

Randomly split the mcycle data into training (75%) and validation (25%) subsets.
```{r}
set.seed(101) # Set Seed so that same sample can be reproduced in future also
# Now Selecting 75% of data as sample from total 'n' rows of the data  
sample <- sample.int(n = nrow(mcycle), size = floor(.75*nrow(mcycle)), replace = F)
train <- mcycle[sample, ]
test  <- mcycle[-sample, ]
```

Using the mcycle data, consider predicting the mean acceleration as a function of time. Use the Nadaraya-Watson method with the k-NN kernel function to create a series of prediction models by varying the tuning parameter over a sequence of values.
```{r}
y <- train$accel
x <- matrix(train$times, length(train$times), 1)

y_test <- test$accel
x_test <- matrix(test$times, length(test$times), 1)
```

```{r}
## k-NN kernel function
## x  - n x p matrix of training inputs
## x0 - 1 x p input where to make prediction
## k  - number of nearest neighbors
kernel_k_nearest_neighbors <- function(x, x0, k=1) {
  ## compute distance betwen each x and x0
  z <- t(t(x) - x0)
  d <- sqrt(rowSums(z*z))
  
  ## initialize kernel weights to zero
  w <- rep(0, length(d))
  
  ## set weight to 1 for k nearest neighbors
  w[order(d)[1:k]] <- 1
  
  return(w)
}
```

```{r}
## Make predictions using the NW method
## y  - n x 1 vector of training outputs
## x  - n x p matrix of training inputs
## x0 - m x p matrix where to make predictions
## kern  - kernel function to use
## ... - arguments to pass to kernel function
nadaraya_watson <- function(y, x, x0, kern, ...) {
  k <- t(apply(x0, 1, function(x0_) {
    k_ <- kern(x, x0_, ...)
    k_/sum(k_)
  }))
  yhat <- drop(k %*% y)
  attr(yhat, 'k') <- k
  return(yhat)
}
```

```{r}
## Compute effective df using NW method
## y  - n x 1 vector of training outputs
## x  - n x p matrix of training inputs
## kern  - kernel function to use
## ... - arguments to pass to kernel function
effective_df <- function(y, x, kern, ...) {
  y_hat <- nadaraya_watson(y, x, x,
                           kern=kern, ...)
  sum(diag(attr(y_hat, 'k')))
}
```

 make predictions using NW method at training inputs
```{r}

y_hat <- nadaraya_watson(y, x, x,kern=kernel_k_nearest_neighbors, k=2)
edf <- effective_df(y, x, 
 kern=kernel_k_nearest_neighbors, k=2)

```
```{r}
y_hat <- nadaraya_watson(y, x, x,kern=kernel_k_nearest_neighbors, k=3)
edf <- effective_df(y, x, 
 kern=kernel_k_nearest_neighbors, k=3)
```

```{r}
y_hat_test <- nadaraya_watson(y, x, x_test,kern=kernel_k_nearest_neighbors, k=3)

```

With the squared-error loss function, compute and plot the training error, AIC, BIC, and validation error (using the validation data) as functions of the tuning parameter.

```{r}
## loss function
## y    - train/test y
## yhat - predictions at train/test x
loss_squared_error <- function(y, yhat)
  (y - yhat)^2

## train error
## y    - train/test y
## yhat - predictions at train x
## loss - loss function
error <- function(y, yhat, loss=loss_squared_error)
  mean(loss(y, yhat))



## AIC
## y    - training y
## yhat - predictions at training x
## d    - effective degrees of freedom
aic <- function(y, yhat, d)
  error(y, yhat) + 2/length(y)*d

## BIC
## y    - training y
## yhat - predictions at training x
## d    - effective degrees of freedom
bic <- function(y, yhat, d)
  error(y, yhat) + log(length(y))/length(y)*d

```

# 
```{r}

train_error<-error(y,y_hat)
test_error<-error(y_test,y_hat_test)
AIC<-aic(y, y_hat, edf)
BIC<-bic(y, y_hat, edf)

## create a grid of inputs 
x_plot <- matrix(seq(min(x),max(x),length.out=100),100,1)

## make predictions using NW method at each of grid points
y_hat_plot <- nadaraya_watson(y, x, x_plot,
                              kern=kernel_k_nearest_neighbors, k=3)

plot(x, y, xlab="Time (ms)", ylab="Acceleration (g)")
legend('topright', legend = c(
paste0('aic = ', round(AIC, 1)),
paste0('bic = ', round(BIC, 1)),
paste0('train_error = ', round(train_error, 1)),
paste0('test_error = ', round(test_error, 1))),
 bty='n')
lines(x_plot, y_hat_plot, col="#882255", lwd=2) 

```

```{r}
library('manipulate')
library('splines') ## 'ns'
library('caret')
```


For each value of the tuning parameter, Perform 5-fold cross-validation using the combined training and validation data. This results in 5 estimates of test error per tuning parameter value.

```{r}

## 5-fold cross-validation of knnreg model
## create five folds
set.seed(1985)
mcycle_flds  <- createFolds(mcycle$accel, k=5)
print(mcycle_flds)
sapply(mcycle_flds, length)  ## not all the same length

cvknnreg <- function(kNN = 10, flds=mcycle_flds) {
  cverr <- rep(NA, length(flds))
  for(tst_idx in 1:length(flds)) { ## for each fold
    
    ## get training and testing data
    mcycle_trn <- mcycle[-flds[[tst_idx]],]
    mcycle_tst <- mcycle[ flds[[tst_idx]],]
    
    ## fit kNN model to training data
    knn_fit <- knnreg(accel ~ times,
                      k=kNN, data=mcycle_trn)
    
    ## compute test error on testing data
    pre_tst <- predict(knn_fit, mcycle_tst)
    cverr[tst_idx] <- mean((mcycle_tst$accel - pre_tst)^2)
  }
  return(cverr)
}
```
Plot the CV-estimated test error (average of the five estimates from each fold) as a function of the tuning parameter. Add vertical line segments to the figure (using the segments function in R) that represent one “standard error” of the CV-estimated test error (standard deviation of the five estimates from each fold).

```{r}
## Compute 5-fold CV for kNN = 1:20
cverrs <- sapply(1:20, cvknnreg)
print(cverrs) ## rows are k-folds (1:5), cols are kNN (1:20)
cverrs_mean <- apply(cverrs, 2, mean)
cverrs_sd   <- apply(cverrs, 2, sd)

## Plot the results of 5-fold CV for kNN = 1:20
plot(x=1:20, y=cverrs_mean, 
     ylim=range(cverrs),
     xlab="'k' in kNN", ylab="CV Estimate of Test Error")
segments(x0=1:20, x1=1:20,
         y0=cverrs_mean-cverrs_sd,
         y1=cverrs_mean+cverrs_sd)
best_idx <- which.min(cverrs_mean)
points(x=best_idx, y=cverrs_mean[best_idx], pch=20)
abline(h=cverrs_mean[best_idx] + cverrs_sd[best_idx], lty=3)
```

erpret the resulting figures and select a suitable value for the tuning parameter.

We can find that with the increase of 'k' in kNN, the CV-estimated test error, to some extent, decreases. When k is 14, the CV-estimated test error is minimal. However, with the increase of 'k' in kNN, the model's bias is increased and its variance is decreased. Thus, for trade_off prinpal, k is 10 that might be a suitable value for the tuning parameter. Because when k is 10, the model complexity is not so high and its test error is small.


