---
title: "Homework 4"
author: Qianyi Sun
date: April 1st, 2020
output: github_document
---

# Load dataset
```{r}
library('ElemStatLearn')
library('randomForest')
```
# keep y column in int type for further calculation
```{r}
vowel.train_int_y<-vowel.train
```
1.Convert the response variable in the “vowel.train” data frame to a factor variable prior to training, so that “randomForest” does classification rather than regression.
```{r}
vowel.train$y
vowel.train$y<- as.factor(vowel.train$y)
str(vowel.train)
```

2.Review the documentation for the “randomForest” function.
3.Fit the random forest model to the vowel data using all of the 11 features using the default values of the tuning parameters.
```{r}
randomForest(y ~ ., data = vowel.train)
```
```{r}
library('manipulate')
library('splines') ## 'ns'
library('caret')
```
4.Use 5-fold CV and tune the model by performing a grid search for the following tuning parameters: 1) the number of variables randomly sampled as candidates at each split; consider values 3, 4, and 5, and 2) the minimum size of terminal nodes; consider a sequence (1, 5, 10, 20, 40, and 80).


```{r}
a<-c(1, 5, 10, 20, 40, 80)
pre_tst<- NA
count <- 0
cverr<- NA
## 5-fold cross-validation of random forest model
## create five folds
set.seed(1985)
vowel.train_flds  <- createFolds(vowel.train$y, k=5)
vowel.train_flds_test  <- createFolds(vowel.train_int_y$y, k=5)
str(vowel.train_flds)
print(vowel.train$y)

cvknnreg <- function(kNN = 10, flds=vowel.train_flds) {
  for(tst_idx in 1:length(flds)) { ## for each fold
    
    ## get training and testing data
    vowel.train_trn <- vowel.train[-flds[[tst_idx]],]
    vowel.train_tst <- vowel.train[ flds[[tst_idx]],]
    vowel.train_tst$y<-as.integer(vowel.train_tst$y)
    
    ## fit random forest model to training data and turn the model
    for (i in 3:5){
      for (j in a ){
        count = count +1
    randomForest_fit <- randomForest(y ~ ., data = vowel.train,mtry = i, nodesize=j)
    
        ## compute test error on testing data
    pre_tst[count] <- predict(randomForest_fit, vowel.train_tst)
    pre_tst[count]<-as.integer(pre_tst[count])
    cverr[count] <- mean((vowel.train_tst$y - pre_tst[count])^2)
      }
    }
  }
  return(cverr)
}
cvknnreg()

```





