---
title: "Homework 2"
author: Qianyi Sun
date: February 5, 2020
output: github_document
---

# Install library
```{r}
library('ElemStatLearn')  ## for 'prostate'
library('splines')        ## for 'bs'
library('dplyr')          ## for 'select', 'filter', and others
library('magrittr')       ## for '%<>%' operator
library('glmnet')         ## for 'glmnet'
```



```{r}
data('prostate')
cor(prostate)
```
## split prostate into testing and training subsets

```{r}

prostate_train <- prostate %>%
  filter(train == TRUE) %>% 
  select(-train)

summary(prostate_train)

prostate_test <- prostate %>%
  filter(train == FALSE) %>% 
  select(-train)
```
# fit the model
```{r}
fit <- lm(lpsa ~ ., data=prostate_train)
```

# Use the testing subset to compute the test error using the fitted least-squares regression model.

```{r}
L2_loss <- function(y, yhat)
  (y-yhat)^2
error <- function(dat, fit, loss=L2_loss)
  mean(loss(dat$lpsa, predict(fit, newdata=dat)))
error(prostate_test, fit)

```

# Train a ridge regression model using the glmnet function, and tune the value of lambda.
```{r}
form  <- lpsa ~ 0 + lweight + age + lbph + lcp + pgg45 + lcavol + svi + gleason
x_inp <- model.matrix(form, data=prostate_train)
y_out <- prostate_train$lpsa
fit <- glmnet(x=x_inp, y=y_out, lambda=seq(0.5, 0, -0.05))
print(fit$beta)

```

# Create a path diagram of the ridge regression analysis
```{r}
plot(x=range(fit$lambda),
     y=range(as.matrix(fit$beta)),
     type='n',
     xlab=expression(lambda),
     ylab='Coefficients')
for(i in 1:nrow(fit$beta)) {
  points(x=fit$lambda, y=fit$beta[i,], pch=19, col='#00000055')
  lines(x=fit$lambda, y=fit$beta[i,], col='#00000055')
}
abline(h=0, lty=3, lwd=2)

```
# Create a figure that shows the training and test error associated with ridge regression as a function of lambda
```{r}
error <- function(dat, fit, lam, form, loss=L2_loss) {
  x_inp <- model.matrix(form, data=dat)
  y_out <- dat$lcavol
  y_hat <- predict(fit, newx=x_inp, s=lam)  ## see predict.elnet
  mean(loss(y_out, y_hat))
}

## compute training and testing errors as function of lambda
err_train_1 <- sapply(fit$lambda, function(lam) 
  error(prostate_train, fit, lam, form))
err_test_1 <- sapply(fit$lambda, function(lam) 
  error(prostate_test, fit, lam, form))

## plot test/train error
plot(x=range(fit$lambda),
     y=range(c(err_train_1, err_test_1)),
     type='n',
     xlab=expression(lambda),
     ylab='train/test error')
points(fit$lambda, err_train_1, pch=19, type='b', col='darkblue')
points(fit$lambda, err_test_1, pch=19, type='b', col='darkred')
legend('topleft', c('train','test'), lty=1, pch=19,
       col=c('darkblue','darkred'), bty='n')

colnames(fit$beta) <- paste('lam =', fit$lambda)
print(fit$beta %>% as.matrix)

```





